---
title: "Bellabeat Case study - Membership conversion"
output: html_notebook
---

# 1. Introduction
Bellabeat is a company that manufactures health-tracking smart products for women, things like the Leaf (wellness tracker), Time (smart watch), Spring (hydration tracker), and the Bellabeat app (activity, sleep, and mindfulness tracking). Founded in 2013 by Urška Sršen and Sando Mur, it positions itself as a lifestyle brand that helps women understand their own health data and build sustainable habits, not just chase step counts.

# 2. Purpose of Analysis
The purpose of this analysis is to understand Bellabeat's current customer behavior and split it into 3 clusters, so the marketing team knows which cluster has the highest chance of converting to a paid membership.

The idea: instead of guessing, we use peer-reviewed research to figure out what kind of user is most likely to stick with a health app long-term. Then we find which of our clusters matches that profile. In plain terms — we find the group of users most likely to actually become paying members, and back it up with published studies rather than gut feeling.

# 3. Data & Methodology

## 3.1 Dataset used
The data source used in this case study is Fitbit Fitness Tracker Data from file Fitabase Data 3.12.16-4.11.16 and merged with Fitabase Data 4.12.16-5.12.16

## 3.2 Information of the dataset
These datasets were generated by respondents to a distributed survey via Amazon Mechanical Turk between 03.12.2016-05.12.2016.  Thirty eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. Individual reports can be parsed by export session ID (column A) or timestamp (column B).  Variation between output represents use of different types of Fitbit trackers and individual tracking behaviors / preferences.  

## 3.3 Preparing
For the packages I use tidyverse for the ggplot2, dplyr, readr, And using factoextra for the K-mean capabilities. 
And for the dataset I merged Fitabase Data 3.12.16-4.11.16 and Fitabase Data 4.12.16-5.12.16 with both being dailyactivity.csv
```{r setup}
library(tidyverse)
library(snakecase)
library(janitor)
library(factoextra)

master_activity_raw <- c(
  
  list.files(path="Fitabase Data 3.12.16-4.11.16", pattern = "dailyActivity", full.names = TRUE),
  list.files(path="Fitabase Data 4.12.16-5.12.16", pattern = "dailyActivity", full.names = TRUE)
) %>% 
  lapply(read_csv) %>% 
  bind_rows()
```

Preview of `master_activity_raw`
```{r preview}
head(master_activity_raw,5)
```

## 3.4 Methodology
- Creating unsupervised cluster segmentation based on criterion that is justified with peer reviewed paper

### 3.4.1 Justification source
- Higher training frequency showed longer engagement  [https://doi.org/10.21037/mhealth-25-15]
- The study suggests that consistency, more than sheer intensity, helps turn tracker use into a long-term habit. [doi:10.2196/22488]
- Intensity level: Moderate - Vigorous / very active will be counted as exercise based on [Activity Levels Explained: Complete Classification Guide. (2026). Free Calculator Hub.]

### 3.4.2 Limitation
- Self selected cohort, all user chose to share data, bias toward already engaged
- tiny sample of 35 users

# 4.1 Processing
Processing that I did included cleaning and analyzing the dataset.

## 4.2 Cleaning
Removing duplicates with distinct and creatinng new column called date and data type Date instead of chr, using if else for safe rerun. And adding weekday
```{r cleaning}
master_activity <- master_activity_raw %>% 
  mutate(
    date = if (is.character(ActivityDate)) mdy(ActivityDate) else ActivityDate,
    weekday = wday(date, label = TRUE, abbr = FALSE)
  ) %>% 
  clean_names() %>% 
  distinct()

# For new column preview
master_activity %>% select(id,date,weekday) %>% head(3)
```
Checking any NA
``` {r, NA}
colSums(is.na(master_activity))
```
No NA's recorded

# 5. Analyzing data
Creating master table and proceeding with K-Mean model

## 5.1 Creating master table
I picked these columns because it will be used as a criterion for upcoming analysis.
```{r, master_table}
master_activity_analysis <- master_activity %>%
  select(
    id, date, weekday,
    total_steps,
    calories,
    very_active_minutes,
    fairly_active_minutes,
    very_active_distance,
    moderately_active_distance,
  )
```

